---
title: "Explore resolution"
author:
  - name: "Tom Smith"
    affiliation: "Cambridge Centre for Proteomics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
abstract: | 
  Here, we explore the resolution at protein and peptide-level
  
output:
  pdf_document:
  html_notebook: default
geometry: margin=1in
fontsize: 11pt
---

Here we explore the resolution for unmodified and modified peptides and unmodified proteins. In a later notebook we will look for modified peptides which have a different classification to the unmodified protein.
```{r}
library(MSnbase)
library(pRoloc)
library(pRolocExt)
library(camprotR)
library(tidyverse)


marker_colours <- getStockcol()[c(1:5, 7:11)]
```

```{r}
unmod_prot <- readRDS(here("results/unmod_prot.rds"))
unmod_pep <- readRDS(here("results/unmod_pep.rds"))
succ_pep <- readRDS(here("results/succ_pep.rds"))

all_data <- list('Protein'=unmod_prot, 'Peptide (unmodified)'=unmod_pep, 'Peptide (succinylated)'=succ_pep)
```


First, let's look at the protein level data for the unmodified proteins as this is the usual level we work from in LOPIT.

```{r}

all_data %>% names() %>% lapply(function(x){
  p <- plotConsProfiles(mrkConsProfiles(all_data[[x]]), plot=FALSE) +
    geom_vline(xintercept=10.5, size=0.25) +
    geom_vline(xintercept=20.5, size=0.25) +
    scale_x_discrete(labels=pData(unmod_prot)$Sample_name) +
    theme(axis.text.x=element_text(size=8, angle=90, vjust=0.5, hjust=1)) +
    ggtitle(x)

})

```

Now, PCA projections
```{r}

PCAres <- function(obj, title){
  plot2D(obj, "markers", method='hexbin', main=title)
  plot2D(obj, "markers", method='hexbin', dims=c(3,4), main=title)

  point_cex <- ifelse(grepl("Peptide", title), 0.5, 1)
  print(point_cex)
  plot2D(obj, "markers", main=title, cex=point_cex, col=marker_colours)
  addLegend(obj, where='bottomright', cex=0.7, col=marker_colours)
  plot2D(obj, "markers", dims=c(3,4), main=title, cex=point_cex, col=marker_colours)
  addLegend(obj, where='bottomright', cex=0.7, col=marker_colours)
}

all_data %>% names() %>% lapply(function(x) PCAres(all_data[[x]], x))

```



Below, we plot the succinylated peptides onto the protein profiles by using the protein PCA rotations.

```{r}
unmod <- all_data$Protein
mod <- all_data$`Peptide (succinylated)`
dims <- c(1,2)
pch <- 20
size <- 1
alpha <- .5

model <- prcomp(exprs(unmod), scale=TRUE, center=TRUE) #  run PCA manually to get rotations
b_pca <- data.frame(predict(model, exprs(mod))) # project modified peptides using umod protein
.pca <- make_proj(unmod, method='PCA', "markers", dims=dims)

cols_labels <- colnames(.pca)[1:2]

.pca <- .pca %>% data.frame()

cols <- colnames(.pca)[1:2]

p <- .pca %>%
  ggplot(aes(!!sym(cols[[1]]), !!sym(cols[[2]]),
             colour=markers,
             alpha=markers=='unknown',
             size=markers=='unknown')) +
  geom_point() +
  theme_camprot(border=FALSE, base_family='sans') +
  xlab(cols_labels[[1]]) +
  ylab(cols_labels[[2]]) +
  scale_colour_manual(values=c(getStockcol()[1:10], colour='grey'), name='') +
  scale_alpha_manual(values=c(1, 0.25), guide=FALSE) +
  scale_size_manual(values=c(1, 0.5), guide=FALSE)

print(p)



p2 <- p + geom_point(data=b_pca,
                     aes_string(sprintf("PC%s", dims[1]), sprintf("PC%s", dims[2])),
                     colour="black", size=size, pch=pch, alpha=alpha)

print(p2)  
```


```{r}

plotModPepOntoUnmodProt <- function(unmod, mod, dims=c(1,2), pch=20, size=1, alpha=.5){

  model <- prcomp(exprs(unmod), scale=TRUE, center=TRUE) #  run PCA manually to get rotations
  b_pca <- data.frame(predict(model, exprs(mod))) # project modified peptides using umod protein
  .pca <- make_proj(unmod, method='PCA', "markers", dims=dims)
  
  cols_labels <- colnames(.pca)[1:2]
  
  .pca <- .pca %>% data.frame()
  
  cols <- colnames(.pca)[1:2]
  
  p <- .pca %>%
    ggplot(aes(!!sym(cols[[1]]), !!sym(cols[[2]]),
               colour=markers,
               alpha=markers=='unknown',
               size=markers=='unknown')) +
    geom_point() +
    theme_camprot(border=FALSE, base_family='sans') +
    xlab(cols_labels[[1]]) +
    ylab(cols_labels[[2]]) +
    scale_colour_manual(values=c(marker_colours, colour='grey'), name='') +
    scale_alpha_manual(values=c(1, 0.25), guide=FALSE) +
    scale_size_manual(values=c(1, 0.5), guide=FALSE)
  
  print(p)
  
  
  
  p2 <- p + geom_point(data=b_pca,
                       aes_string(sprintf("PC%s", dims[1]), sprintf("PC%s", dims[2])),
                       colour="black", size=size, pch=pch, alpha=alpha)
  
  print(p2) 
    
  invisible(list("only_umod"=p, "inc_mod"=p2))
}


projected_onto_protein <- plotModPepOntoUnmodProt(all_data$Protein, all_data$`Peptide (succinylated)`, size=0.5)

projected_onto_peptide <- plotModPepOntoUnmodProt(all_data$`Peptide (unmodified)`, all_data$`Peptide (succinylated)`, size=0.5)
ggsave(here("figures/peptide_pca.png"), projected_onto_peptide$inc_mod)
ggsave(here("figures/peptide_pca.pdf"), projected_onto_peptide$inc_mod)


```

