---
title: "Classify major localisation"
author:
  - name: "Tom Smith"
    affiliation: "Cambridge Centre for Proteomics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
abstract: | 
  Here we classify proteins/peptides according to their marker profiles in the
  unenriched/unmodified LOPIT, using SVM
output:
  pdf_document:
  html_notebook: default
geometry: margin=1in
fontsize: 11pt
---



```{r}
library(MSnbase)
library(pRoloc)
library(pRolocExt)
library(camprotR)
library(tidyverse)


marker_colours <- getStockcol()[c(1:5, 7:11)]
```

```{r}
unmod_prot <- readRDS(here("results/unmod_prot.rds"))
unmod_pep <- readRDS(here("results/unmod_pep.rds"))
succ_pep <- readRDS(here("results/succ_pep.rds"))

```




SVM parameter optimisation. Not re-run by default to avoid excessive run-time.
```{r}
times <- NULL
times$prot <- 50
times$pep <- 30

```

We are using just the unmodified peptides here although we could use the concatenated datasets we create below since all markers for the modified peptides are set to 'unknown'.
```{r, eval=FALSE}

sigma <- 2^(-5:0)
cost <- 2^(-2:4)

# less onerous for quicker check
#times <- 10
#sigma <- 2^(-4:-1)
#cost <- 2^(0:3)

params_unmod_prot <- pRoloc::svmOptimisation(
  unmod_prot,
  fcol="markers",
  times=times$prot,
  cost=cost,
  sigma=sigma)




saveRDS(params_unmod_prot, here("results/svm_params_unmod_prot.Rds"))

plot(params_unmod_prot)
levelPlot(params_unmod_prot)
getParams(params_unmod_prot)


```

For the optimisation of peptide-level SVM parameters, we will use the protein-level optimal parameters as a start point and vary 2-fold in either direction. We will also limit to 30 iterations. This is to limit the run-time of the optimisation.
```{r, eval=FALSE}

sigma <- 2^(-5:-3)
cost <- 2^(2:4)

params_unmod_pep <- pRoloc::svmOptimisation(
  unmod_pep,
  fcol="markers",
  times=times$pep,
  cost=cost,
  sigma=sigma)

saveRDS(params_unmod_pep, here("results/svm_params_unmod_pep.Rds"))
```


Read in previous SVM parameter optimisation
```{r}
params_unmod_prot <- readRDS("../results/svm_params_unmod_prot.Rds")
params_unmod_pep <- readRDS("../results/svm_params_unmod_pep.Rds")

```

Now let's inspect the marker set level F1 scores
```{r}
svm_params <- NULL
svm_params$prot <- params_unmod_prot
svm_params$pep <- params_unmod_pep


svm_params %>% names() %>% lapply(function(condition){
    data.frame(getF1Scores(svm_params[[condition]])) %>%
      mutate(condition=condition)
  }) %>% do.call(what='rbind') %>%
      ggplot(aes(condition, F1)) +
    geom_boxplot(notch=TRUE) +
    theme_camprot() +
    ylim(0.7, 1) +
    xlab('') +
    ylab('Macro F1') +
  scale_colour_manual(values=get_cat_palette(3), name='')
```

OK, so very little difference. Protein macro F1 is higher as we might expect but not a huge difference

What about the organelle level F1
```{r}

organalle_f1 <- svm_params %>% names() %>% lapply(function(condition){
  
  params <- svm_params[[condition]]
  
  f1scoreperO <- matrix(0, times[[condition]], dim(params@cmMatrices[[1]])[1])
  
  for(i in 1:times[[condition]]){
    
    conf <- params@cmMatrices[[i]]
    
    f1perO <- MLInterfaces::F1(conf, naAs0 = TRUE)
    
    f1scoreperO[i, ] <- f1perO
  }
  
  OrganelleF1 <- as.data.frame(f1scoreperO)
  colnames(OrganelleF1) <- getMarkerClasses(unmod_pep)
  
  OrganelleF1 %>% mutate(condition=condition) 
}) %>% do.call(what='rbind')


organalle_f1 %>% pivot_longer(-condition, names_to='marker_set', values_to='F1') %>%
  ggplot(aes(marker_set, F1, colour=condition)) +
  geom_boxplot(position='dodge', notch=TRUE) +
  theme_camprot(base_size=12) +
  scale_colour_manual(values=get_cat_palette(4)[3:4], name='') +
  theme(strip.text=element_text(size=8), axis.text.x=element_text(angle=45, hjust=1, vjust=1), aspect.ratio=0.2) +
  xlab('')
```  

Similar story. In fact, nucleus F1 is sometimes better with peptide!


Combine modified and unmodified peptides so we can use the markers from the unmodified LOPIT and the optimised SVM parameters for these markers to classify all peptides together.
```{r}
combineLOPITs <- function(lopit1, lopit2){
  
  tmp_exprs <- rbind(exprs(lopit1), exprs(lopit2))
  
  tmp_fData1 <- fData(lopit1)
  tmp_fData2 <- fData(lopit2)
  
  for(col in setdiff(colnames(tmp_fData2), colnames(tmp_fData1))){
    tmp_fData1[[col]] <- NA
  }
  
  for(col in setdiff(colnames(tmp_fData1), colnames(tmp_fData2))){
    tmp_fData2[[col]] <- NA
  }
  
  tmp_fData <- rbind(tmp_fData1, tmp_fData2)
  
  mod_prot_plus_mod_pep <- MSnSet(tmp_exprs, tmp_fData, pData(lopit1))
  
  invisible(mod_prot_plus_mod_pep)
}

combineUnmodProtModPep <- function(unmod_prot, mod_pep, mod_to_unknown=FALSE){
  
  combined_data <- combineLOPITs(unmod_prot, mod_pep)
  fData(combined_data)$mod <- c(rep("Unmodified", length(rownames(unmod_prot))),
                                rep("Modified", length(rownames(mod_pep))))
  
  for(col in fvarLabels(combined_data)[grep("^CV", fvarLabels(combined_data))])
    
    if(mod_to_unknown){
      # reset the marker column to unknown so that all modified features can be classified  
      fData(combined_data)$markers[fData(combined_data)$mod=="Modified"] <- "unknown"
    }
  
  invisible(combined_data)
}

unmod_pep_succ_pep <- combineUnmodProtModPep(unmod_pep, succ_pep, mod_to_unknown=TRUE)


# Just checking the concatenated dataset is the right dimensions
dim(unmod_pep_succ_pep)
dim(unmod_pep)
dim(succ_pep)
```
Crude function to filter SVM classifications by a quantile threshold.
```{r}
addFilteredSVM <- function(obj, quant_thresh=0.75){
  fData(obj)$max.svm <- apply(fData(obj)$svm.all.scores, MARGIN=1, FUN=max)
  
  obj_no_markers <- obj[fData(obj)$markers=="unknown",]

  svm_individual_threshold <- fData(obj_no_markers) %>%
    group_by(svm) %>%
    summarise(threshold=quantile(max.svm, quant_thresh))
    
  p <- fData(obj_no_markers) %>%
    data.frame() %>%
    ggplot() +
    geom_boxplot(aes(svm, max.svm)) +
    theme_camprot() +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1, size=10))
  
  print(p)
  
  print(svm_individual_threshold)
  svm_individual_threshold_list <- svm_individual_threshold$threshold
  names(svm_individual_threshold_list) <- svm_individual_threshold$svm
  
  fData(obj) <- fData(obj) %>%
    tibble::rownames_to_column() %>%
    rowwise() %>%
    mutate("svm_filtered_ind_thresh"=ifelse(max.svm<svm_individual_threshold_list[[svm]], "unknown", as.character(svm))) %>%
    tibble::column_to_rownames()
  
  print(table(fData(obj)$svm_filtered_ind_thresh))
  print(fData(obj) %>% filter(markers=="unknown") %>%
          pull(svm_filtered_ind_thresh) %>% table())
  
  fData(obj)$svm_filtered_single_threshold <- as.character(fData(obj)$svm)
  low_confidence <- apply(fData(obj)$svm.all.scores, MARGIN=1, FUN=max)<0.9
  fData(obj)$svm_filtered_single_threshold[low_confidence] <- "unknown"
  
  print(table(fData(obj)$svm_filtered_single_threshold))
  print(fData(obj) %>% filter(markers=="unknown") %>%
          pull(svm_filtered_single_threshold) %>% table())
  
  invisible(obj)
}


```

Classify peptides using SVM
```{r}

best_params <- getParams(params_unmod_pep)

pep_classifications <- svmClassification(
    unmod_pep_succ_pep, scores="all", cost=best_params[['cost']],
    sigma=best_params[['sigma']], fcol="markers")


pep_classifications <- pep_classifications %>% addFilteredSVM()


saveRDS(pep_classifications, here("results/pep_classifications.rds"))
```


